---
title: "CourseMatch Analytics - Spring 2019"
subtitle: "Data Processing for Google Sheet"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  bookdown::html_document2:
    #css: css/analysis.css
    code_folding: show
    df_print: paged
    number_sections: yes
    self-contained: true
    fig_caption: yes
    toc: true
    toc_float: true
params:
  external_queries: false
---

```{r setup, include = TRUE}
#setwd("<insert path>")

# knitr options
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center',
                      fig.width = 9, fig.height = 8, cache = FALSE)

# Forces any incline R to only have 2 digits and a comma for the thounsands place
knitr::knit_hooks$set(inline = function(x) {if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })

# Don't show scientific notation and only 3 digits
options(scipen = 999, digits = 3)

# Load packages
pacman::p_load(tidyverse, readxl, stringr, lubridate, googlesheets, rvest, forcats, 
               grid, gridExtra, htmlTable, highcharter)

# Token to connect to Googlespreadsheets
#token <- gs_auth(cache = FALSE)
#saveRDS(token, file = "googlesheets_token.rds")
suppressMessages(gs_auth(token = "googlesheets_token.rds", verbose = FALSE))

# Create corresponding folders in working directory
data_path <- "data/"; viz_path <- "viz/"

#Set ggplot theme
ggplot2::theme_set(theme_minimal() + 
                     theme(legend.position="top", plot.caption = element_text(size = 8), 
                           panel.grid.minor.x = element_blank(), 
                           strip.background = element_rect(fill = "grey85", colour = "grey85"),
                           legend.margin=margin(t = -0.3, r = 0, b = -0.4, l=0, unit="cm")))

wharton_col <- c("#004785", "#A90533", "#A262460", "#282F85", "#026CB5", "#06AAFC",
                 "#532A85","#96227D", "#A8204E","#C5093B","#2D2C41","#D7BC6A",
                 "#B2B6A7","EEEDEA")
names(wharton_col) <- c("wharton_blue", 'wharton_red','midnight_blue','atlantic_blue','pacific_blue','bay_blue',
                        'evening_red','sunset_red','brick_red','heritage_red','night_street','morning_yellow',
                        'college_grey','marine_grey')

hcoptslang <- getOption("highcharter.lang")
hcoptslang$thousandsSep <- ","
hcoptslang$numericSymbols <- c("k", "M", "B", "T", "P", "E")
options(highcharter.lang = hcoptslang)
```

## Summary

1. This notebook documents how we built the [Google Sheet](https://docs.google.com/spreadsheets/d/18PgeLV0QIaRghlBXfa_LXyYXMVbHqY2AqyaTB4teZm0/edit?usp=sharing) with course and instructor evaluations and clearing prices
2. There are some basic exploratory plots at the end

# Data Acquisition

## CourseMatch Sections

To get all sections listed in CourseMatch, we extract the table from the html. First, save the home CourseMatch interface page as a .html file in the data folder. Some courses are cross-listed and we create a column to include the other sections a course can be listed as. 

```{r parse-coursematch-html}
# List of sections available for selection within CourseMatch
coursematch_sections <- 
  read_html(paste0(data_path, "Course Match.html")) %>%
  html_node(xpath = '//*[@id="preferences-table"]') %>%
  html_table(header = TRUE, trim = TRUE) %>%
  as_tibble() %>%
  select(1) %>%
  rename_at(vars(contains("Dept")), funs(paste0("Course_Multiple"))) %>%
  mutate(Section = str_extract_all(Course_Multiple, "[A-Z]{4}[0-9]{6}")) %>%
  unnest() %>%
  mutate(cross_listed_as = str_replace(Course_Multiple, Section, "")) %>%
  mutate(cross_listed_as = if_else(cross_listed_as == "", NA_character_, cross_listed_as)) %>%
  distinct(Section, cross_listed_as) %>%
  rowwise() %>%
  mutate(cross_listed_as = if_else(cross_listed_as == "", NA_character_, 
                                   paste0(str_extract_all(cross_listed_as, 
                                                          "[A-Z]{4}[0-9]{6}", simplify = TRUE), collapse = ","))) %>%
  ungroup()
```


## Course Information

To get all the information about the Fall 2017 courses, we can use the *export* feature on [Spike](https://spike.wharton.upenn.edu/courses/index.cfm?method=search_courses). Unfortunately, this contains non-MBA classes too (ungrad and PhD).

```{r load-all-wharton-courses}
spring_2019_courses <- 
  readxl::read_excel(paste0(data_path, "export_courses.xls"))
```


## Course Evaluations

Course evaluation comes from a [Spike export](https://spike.wharton.upenn.edu/courses/?method=export_evaluations). You need to open the Excel file and resave before it will load into R (see documented issue caused by the way this file was built by Spike). 

```{r load-course-eval-data}
course_eval <- 
  readxl::read_excel(paste0(data_path, "CourseEvaluationData.xls"), skip = 1) %>% 
  filter(n() != row_number())
```


## Clearing Prices

Clearing price data comes the **public** [MBA Inside website](https://mba-inside.wharton.upenn.edu/course-match/). We save the files corresponding to the last 4 semesters, from Spring 2017 to Fall 2018

```{r download-clearing-prices}

# Function to load each xlsx file of clearing prices
# Need to have a switch because the Fall 2013 file
# contains an extra row at the top
fn_load_clearing_price_xlsx <- function(file_name) {
  if (str_detect(file_name, "13")) {
    readxl::read_excel(file_name, skip = 1)
  } else {
    readxl::read_excel(file_name, skip = 0)
  }
}

# Function to rename the columns that 
# look like 'section' and 'price' to Section
# and Price so that they can be binded
fn_rename_at <- function(df, string) {
  df %>%
    rename_at(vars(matches(string)), funs(paste0(string)))
}

# Load and coalesce each of the clearing price
# Excel files
clearing_prices <- 
  list.files(paste0(data_path, "clearing_prices/"), full.names = TRUE) %>%
  map(fn_load_clearing_price_xlsx) %>%
  map(select, matches("Section|Price")) %>%
  map(fn_rename_at, string = "Section") %>%
  map(fn_rename_at, string = "Price") %>%
  bind_rows() %>%
  mutate(Course = str_sub(Section, 1, 7)) %>%
  select(Course, Section, Price)
```

# Data Cleaning

After acquiring data from the four data sources we need to do some combining and cleaning.

```{r spring_2019}
spring_2019 <- 
  spring_2019_courses %>%
  select(-Max, -Status) %>%
  distinct() %>%
  mutate(Section = strsplit(as.character(Section), " / ")) %>% 
  unnest(Section) %>%
  select(Section, everything()) %>%
  inner_join(coursematch_sections, by = "Section") %>%
  mutate(title_length = length(Title)) %>%
  arrange(Section, desc(title_length)) %>%
  group_by(Section) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  mutate(
    Course = str_sub(Section, 1, 7)
    , Section = str_sub(Section, 8, 10)
    , CU = as.numeric(CU)
  ) %>%
  select(Course, Section, Title, Instructor, CU, Dates, `Day(s)`, cross_listed_as) %>%
  separate(Dates, c("start_date", "end_date"), sep = "-") %>%
  mutate(
    start_date = lubridate::mdy(start_date)
    , Days = stringr::str_extract(`Day(s)`, "^([\\w]+)")
    , Quarter = if_else(CU == 1, "Full", if_else(lubridate::month(start_date) == 8, "Q1", "Q2"))
    , start_time = stringr::str_extract(`Day(s)`, "(?:\\S+){2}(\\S+)")
    , end_time = str_trim(stringr::str_extract(`Day(s)`, "\\s(\\S+)$"))
    , Title = str_to_title(Title)
  ) %>%
  select(-start_date, -end_date, -`Day(s)`)
```

```{r tbl_questions}
tbl_questions <- tribble(
  ~question_type, ~question, ~question_abbr
  , 'Course', 'Overall Quality of Course (0=worst 4=best)', 'Course Quality'
  , 'Course', 'Value of Assigned Readings (0=worst 4=best)', 'Value of Assigned Readings'
  , 'Course', 'Learned from this Course in terms of Knowledge / Concepts / Skills / Thinking Ability (0=worst 4=best)', 'Amount Learned'
  , 'Course', 'Rate the Difficulty of this Course (0=easiest 4=hardest)', 'Difficulty'
  , 'Course', 'Rate the Amount of Work Required for this Course (0=easiest 4=hardest)', 'Amount of Work Required'
  , 'Course', 'Would you Recommend this Course to a Major? (4=most strongly)', 'Recommend to Major?'
  , 'Course', 'Would you Recommend this Course to a Non-Major? (4=most strongly)', 'Recommend to Non-Major'
  , 'Instructor', 'Overall Quality of Instructor (0=worst 4=best)', 'Instructor Quality'
  , 'Instructor', 'Instructor Ability to Communicate Subject Matter (0=worst 4=best)', 'Ability to Communicate Subject Matter'
  , 'Instructor',  'Instructor Ability to Stimulate Student Interest (0=worst 4=best)', 'Ability to Stimulate Interest'
  , 'Instructor', 'Instructor Accessibility / Willingness to Discuss Course Content / Problems (0=worst 4=best)', 'Instructor Accessibility'  
) %>%
  mutate(
    question_type = fct_inorder(question_type)
    , question_abbr = fct_inorder(question_abbr)
  )
```

```{r tbl_course_lookup}
tbl_course_lookup <-
  tribble(
    ~evaluated_course, ~coursematch_course,
    'ACCT208',	'ACCT718',
    'BEPP201',	'BEPP770',
    'BEPP305',	'BEPP805',
    'STAT451',	'BEPP851',
    'STAT453',	'BEPP853',
    #'REAL721',	'FNCE721',
    #'OIDD693',	'LGST693',
    #'LGST809',	'MGMT815',
    'OIDD415',	'OIDD515',
    'STAT470',	'STAT770',
    'STAT451',	'STAT851',
    'STAT453',	'STAT853',
    NA_character_, 'MGMT833',
    NA_character_, 'LGST641',
    'OIDD691', 'MGMT691', 
    'LGST806', 'MGMT691',
    'BEPP289', 'BEPP789',
    'OIDD261', 'BEPP761',
    'LGST202', 'LGST802',
    'LGST220', 'LGST820',
    'LGST215', 'LGST815',
    'MKTG224', 'MKTG724',
    'OIDD397', 'OIDD697',
    'OIDD261', 'OIDD761',
    'REAL215', 'REAL724',
    'MKTG212', 'MKTG712'
  )
```

```{r cross_listed_course}
cross_listed_course <- 
  tribble(
    ~coursematch, ~evaluated_as
      , "BEPP708", "REAL708" # Housing Markets
      , "OIDD763", "BEPP763" # Energy Markets and Policy
      , "FNCE721", "REAL721" # Real Estate Investments
      , "LGST692", "OIDD692" # Advanced Topics in Negotiations
      , "LGST804", "REAL804" # Real Estate Law
      , "LGST806", "MGMT691" # Negotiations 
      , "MGMT815", "LGST809" # Sports Business Mgmt
      , "MGMT692", "OIDD692" # Advanced Topics in Negotiations        
      , "STAT776", "MKTG776" # Applied Probablistic Models in Marketing
      , "OIDD691", "MGMT691" # Negotiations 
      , "MGMT690", "OIDD690" # Managerial Decisn Making
      , "LGST693", "OIDD693" # Influence
  )
```

```{r course_eval_cleaned}
course_eval_cleaned <- 
  course_eval %>%
  gather(question, value, -Term, -Section, -Title, -Instructor) %>%
  left_join(tbl_questions, by = 'question') %>%
  select(-question) %>%
  mutate(
    Year = str_sub(Term, 1,4)
    , Course = str_sub(Section, 1, 7)
  ) %>%
  left_join(tbl_course_lookup, by = c('Course' = 'evaluated_course')) %>%
  mutate(Course = if_else(is.na(coursematch_course), Course, coursematch_course)) %>%
  mutate(
    Course_Num = str_sub(Course, 5, 7)
    , Dept = str_sub(Course, 1, 4)
  ) %>%
  select(Year, Dept, Course_Num, Course, Section, Title, Instructor, question_type, question_abbr, value) %>%
  # OIDD515 (Product Design) is the only CourseMatch class with a course number less than 600
  filter(Course == "OIDD515" | (Course_Num >= 600 & Course_Num < 900))   
```

```{r show_course_eval_cleaned, echo = FALSE}
course_eval_cleaned %>% 
  sample_n(10) %>% 
  htmlTable(
    rnames = FALSE
    , caption = "10 sample rows from the tidied course evaluation data"
    , css.cell = "padding-left: .5em; padding-right: .5em;"
    , header = c('Year','Dept','Course Number','Course','Section','Title','Instructor','Question Type', 'Question', "Value")
    , align = c("cccccllclr")
    , col.rgroup = c("none", "#F7F7F7")
  )
```

# Summaries for Google Sheet

We present that data at two levels of detail:

1. **Course View**: Aggregate the reviews for all sections a course was taught
2. **Instructor View**: Aggregate the reviews for all sections an instructor has taught

We also create a "benchmark" that uses all historical data (not just the courses that will be taught in Spring 2019).

## Course View

```{r course_metrics}
course_metrics <- 
  spring_2019 %>%
  left_join(cross_listed_course, by = c('Course' = 'coursematch')) %>% 
  mutate(evaluated_as = if_else(is.na(evaluated_as), Course, evaluated_as)) %>%
  left_join(
    course_eval_cleaned %>%
      filter(question_type == "Course") %>%
      group_by(Course, question_abbr) %>%
      summarise(
        value = mean(value)
        , sections_evaluated = n()
      ) %>%
      spread(question_abbr, value)
    , by = c('evaluated_as' = 'Course')
  ) %>%
  left_join(
    clearing_prices %>%
      filter(str_sub(Course, 1, 4) != "WHCP") %>%
      mutate(Course = str_replace(Course, "OPIM", "OIDD")) %>%
      left_join(cross_listed_course, by = c('Course' = 'coursematch')) %>%
      mutate(evaluated_as = if_else(is.na(evaluated_as), Course, evaluated_as)) %>%
      group_by(evaluated_as) %>%
      summarise(clearing_price = mean(Price)) %>%
      ungroup() %>%
      mutate(percentile = percent_rank(clearing_price))
    , by = c("evaluated_as")
  ) %>%
  select(-evaluated_as) %>%
  select(-one_of('cross_listed_as'), one_of('cross_listed_as'))
```

```{r course_benchmark}
course_benchmark <- 
  course_eval_cleaned %>%
  filter(question_type == "Course") %>%
  group_by(question_abbr) %>%
  summarise(
    value = mean(value)
    , sections_evaluated = n()
  ) %>%
  spread(question_abbr, value) %>%
  bind_cols(
    clearing_prices %>%
      filter(str_sub(Course, 1, 4) != "WHCP") %>%
      summarise(clearing_price = mean(Price))
  )
```

## Instructor View

```{r instructor_metrics}
instructor_metrics <- 
  spring_2019 %>%
  mutate(instructors = strsplit(Instructor, "/")) %>% 
  unnest(instructors) %>%
  mutate(instructors = str_trim(instructors)) %>%
  filter(!(is.na(instructors) | instructors == "")) %>%
  distinct(instructors) %>%
  left_join(
    course_eval_cleaned %>%
      filter(question_type == "Instructor") %>%
      group_by(Instructor, question_abbr) %>%
      summarise(
        value = mean(value)
        , sections_evaluated = n()
      ) %>%
      spread(question_abbr, value)
    , by = c('instructors' = 'Instructor')
  ) %>%
  arrange(instructors)
```

```{r instructor_benchmark}
instructor_benchmark <- 
  course_eval_cleaned %>%
  filter(question_type == "Instructor") %>%
  group_by(question_abbr) %>%
  summarise(
    value = mean(value)
    , sections_evaluated = n()
  ) %>%
  spread(question_abbr, value)
```

## Export to Google Sheet

We then use the [googlesheets](https://cran.r-project.org/web/packages/googlesheets/) package to add them to the [Google Sheet](https://docs.google.com/spreadsheets/d/18PgeLV0QIaRghlBXfa_LXyYXMVbHqY2AqyaTB4teZm0/edit?usp=sharing)


```{r update_google_sheet}
if (params$external_queries) {
  gs <- gs_title("Analytics Club - CourseMatch Spring 2019 (File --> Make a copy)")
    
  # Course View tab
  gs %>% gs_edit_cells(ws = "Course View", input = course_benchmark, anchor = "K3", col_names = FALSE)
  gs %>% gs_edit_cells(ws = "Course View", input = course_metrics, anchor = "B4", col_names = FALSE)
    
  # Instructor View tab
  gs %>% gs_edit_cells(ws = "Instructor View", input = instructor_benchmark, anchor = "C3", col_names = FALSE)
  gs %>% gs_edit_cells(ws = "Instructor View", input = instructor_metrics, anchor = "B4", col_names = FALSE)
}
```
